---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nbbSTSestimate

<!-- badges: start -->
<!-- badges: end -->

Multiprocessing time series imputation by State Space models (incl. low frequency input)


## Description
Performs in-sample imputation of missing values, as well as for-now- and back-casting of univariate time series by integrating  (or not) low-frequency input.


## Installation

```{r, echo = TRUE, eval = FALSE}
remotes::install_github("clemasso/nbbSTSestimate")
```


## Usage (2 possibilites)

### 1) Input in R

```{r, echo = TRUE, eval = TRUE}
library("nbbSTSestimate")

# Examples
## Input: retail data, 3 series
retail_extract <- cbind(
    GroceryStores = rjd3toolkit::retail$GroceryStores,
    BookStores = rjd3toolkit::retail$BookStores,
    ShoeStores = rjd3toolkit::retail$ShoeStores
)
x <- window(retail_extract, c(1990, 1), c(2012, 6), extend = TRUE)
x.lf <- ts(rbind(
    matrix(c(325000, 330000, 7800, 8000, 17500, 18000), nrow = 2, byrow = FALSE),
    aggregate.ts(retail_extract, nfreq = 1),
    matrix(c(525000, 15000, 27000), nrow = 1)
), start = c(1990, 1), frequency = 1)
## Estimate
rslt1 <- estimateSTS(x, x.lf) # auto procedures for model selection, outlier detection and calendar effect by default
plot(rslt1,"BookStores")
```

```{r, echo = TRUE, eval = TRUE}
library("nbbSTSestimate")

## Adding specific input
outliers_input <- data.frame(
    series_name = c("GroceryStores", "GroceryStores", "BookStores", "ShoeStores"),
    outliers = c("LS_2008M12", "AO_2002M04", "auto", "auto")
)
rslt2 <- estimateSTS(
    x,
    x.lf,
    stsmodel = "bsm",
    outliers = "userdefined",
    outliers.df = outliers_input,
    cal.effect = "auto",
    cal.effect.td = "BE",
    cal.effect.easter = TRUE
)
plot(rslt2,"BookStores")
```

### 2) Input in Excel

```{r, echo = TRUE, eval = FALSE}
# Examples
## Fill structured .xlsx file with your input. See template or vignette for the structure of the Excel file
res <- estimateSTS_fromXLSX(
    path.data = "input_to_R.xlsx",
    is.lf = TRUE,
    conversion = "Sum",
    path.output = "output_R.xlsx"
)
```


## Contributing

Any contribution is welcome and should be done through pull requests and/or issues.

## Licensing

The code of this project is licensed under the [European Union Public Licence (EUPL)](https://joinup.ec.europa.eu/page/eupl-text-11-12).
